### 2016-07-18 article/article.tex
+ assorted ms fixes including Rich's edits + results analysed + change to raptr

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rapr: Representative and Adequate Prioritisations in R},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{rapr: Representative and Adequate Prioritisations in R}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$,
Richard A. Fuller$^1$\\$^1$School of Biological Sciences, The University
of Queensland, Brisbane, QLD, Australia\\$^2$School of Geography,
Planning and Environmental Management, The University of Queensland,
Brisbane, QLD, Australia\\Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{18 July 2016}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,titling,natbib,pdfpages,lineno,booktabs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% format toc
%\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

% line numbers
\linenumbers

% format abstract
\renewcommand{\abstractname}{Summary}
\renewenvironment{abstract}
 {\small
  \begin{center}
  \bfseries \abstractname\vspace{-.5em}\vspace{0pt}
  \end{center}
  \list{} {%
   \setlength{\leftmargin}{2mm}
   \setlength{\rightmargin}{\leftmargin}%
  }%
  \item\relax}
{\endlist}

% format section headers
\titleformat*{\section}{\Large\bfseries}
\titleformat{\subsection}[display]
	{\large\sffamily\lsstyle}
	{\subsectiontitlename\ \thesubsection}
	{0.5em}{}
\titleformat*{\subsubsection}{\large\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

% define struts for tables
\newcommand\T{\rule{0pt}{2.6ex}} % top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut

% define command to put new lines in table cells
%\newcommand{\specialcell}[2][c]{%
%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\begin{document}

\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  An underlying aim in spatial conservation planning is to maximise the
  long-term persistence of biodiversity. Yet current approaches
  overwhelming focus on ``adequacy'', the concept ensuring a target
  proportion of the distribution of a biodiversity feature (species or
  ecosystem) is reserved. Ultimately, to achieve long-term persistence,
  the processes the sustain biodiversity must also be conserved. Many
  biodiversity processes can be preserved by conserving a representative
  sample of each biodiversity feature. This idea has for decades been
  encapsulated in principle of ``representativeness''. For example,
  representing the full range of genetic diversity or ability to persist
  in a range of climatic conditions might be crucial for the long -term
  persistence of a species in a world where environmental change is
  accelerating. However, fully operationalising the principle of
  repressiveness in spatial conservation planning has proven
  methodologically challenging.
\item
  To address this issue, we developed the
  \texttt{rapr: Representative and Adequate Pioritisations in R}, a
  toolkit to guide reserve selection using explicit targets for
  representing biodiversity processes as well as meeting area protection
  targets. Using a novel formulation of the reserve selection problem,
  users set ``space targets'' to secure a representative sample of each
  species within an attribute space reflecting any biodiversity process
  (eg. an attribute space expressing variation in genetic
  characteristics between individuals in an area, or variation in
  climatic conditions between areas occupied by indivduals). We explored
  the functionality of this \texttt{R} package using a simulation study
  and two case studies. In each study, we generated prioritisations that
  aimed to preserve an amount of habitat for each species using amount
  target--to represent conventional reserve selection methods--and
  compared them with prioritisations generated using both habitat and
  space targets--to represent this method.
\item
  We show that explicitly considering biodiversity processes in reserve
  selection can substantially change the configuration of the resulting
  solution. Simulations suggest that including representativeness in
  conservation planning is particularly important where the biodiversity
  feature has multi-model distributions in an attribute space. Results
  from initial case studies show that explicitly considering space-based
  targets for biodiversity processes can result in prioritisations that
  do not necessarily require much more area than traditional approaches,
  but that could be much more effective in achieving long term
  biodiversity persistence.
\item
  The \texttt{rapr R} package provides a unified framework achieving
  spatial conservation prioritisations that are both adequate and
  representative, resulting in a greater chance of long term
  biodiversity persistence that area-based planning alone.
\end{enumerate}
\end{abstract}

Short running title: Representative and adequate prioritisations
\newline
Abstract word count: 370 / 350\\Total word count: XXXXX / 8000\\Number
of references: 34\\Number of figures: 8\\Number of tables: 0\\Number of
textboxes: 0\\Number of equations: 8\\\newpage

\section{Introduction}\label{introduction}

Perhaps the most fundamental aim of conservation is to maximise the
long-term persistence of biodiversity at organisational levels (McNeely
1994; Margules \& Pressey 2000). To achieve this, conservation actions
must preserve biodiversity patterns (eg. populations, species,
ecosystems), but also crucially the processes that sustain them. One of
the major tangible achievements of modern conservation has been the act
of setting aside areas for preservation (Sanderson \emph{et al.} 2015).
Reserve networks buffer species from gross threatening processes and set
the stage for enhanced management interventions (Gaston \emph{et al.}
\{2008\}). However, the resources available for conservation action are
limited, and so reserve networks must be sited in places that satisfy
conservation objectives for minimum cost (Margules \& Pressey 2000). To
achieve this, reserve selection is often formulated as an optimisation
problem and then solved to identify cost-effective candidate reserve
systems (termed prioritisations; Margules \& Pressey 2000). Many
decision support tools have been developed to help identify
cost-effective solutions (eg. \texttt{ ConsNet}, Ciarleglio \emph{et
al.} 2009;\texttt{ Marxan}, Ball \emph{et al.} 2009;\texttt{ Zonation},
Moilanen 2007). Typically, these tools are used to deliver a
prioritisation that preserves a set of species of interest to the
conservation planner (termed target species or features). Given a set of
species, decision makers can use these tools to preserve biodiversity
patterns by generating a prioritisation that secures an adequate
proportion of each species' range.

However, to achieve effective conservation, reserve networks must do
more than cover enough area. They must capture the processes that
sustain biodiversity over the long term; they must capture both
ecological and evolutionary processes (Crandall \emph{et al.} 2000;
Margules \& Pressey 2000). Ecological processes, such as predator-prey
interactions, pollination, and decomposition, are required for
biodiversity to persist over short time-scales. Typically, they operate
over small geographic domains--with exceptions such as migration and
refugial habitats--and can be preserved using suitably large planning
units (Ciarleglio \emph{et al.} 2009) that each contain a discrete unit
of habitat (Klein \emph{et al.} 2009). On the other hand, evolutionary
processes are required for biodiversity to persist over long
time-scales, and they typically operate over large geographic domains.
Protected areas must represent adaptive evolutionary processes to foster
resilience against environmental change (eg. climate change; Pyke \&
Fischer 2005). Protected areas must also preserve neutral evolutionary
processes (Moritz 1994, 1999, 2002), arising from restriction of gene
flow between populations. They are important for maintaining genetic
diversity, and avoiding inbreeding depression. In recent decades, a
wealth of data relevant to evolutionary processes has become available
to conservation planners (eg. bioclimatic, genetic, and trait data;
Hijmans \emph{et al.} 2005; Raymond \emph{et al.} 2015; Jones \emph{et
al.} 2009). Yet such data are rarely used to guide conservation planning
(Hendry \emph{et al.} 2010). This is perhaps in large part due to that
fact that existing reserve selection tools focus on capturing either
biodiversity patterns or processes--but not both.

Existing conservation planning tools are generally not well-suited for
representing biodiversity processes (but see Faith 2003). To preserve
biodiversity processes, a prioritisation must capture a representative
sample of each species. For instance, to preserve predator-prey
interactions, a prioritisation must preserve individuals from each
predator and prey species in the same area. To preserve adaptive
evolutionary processes, a prioritisation must preserve the adaptive
landscape of each species--populations experiencing different selection
pressures (Moritz 2002). To preserve neutral evolutionary processes, a
prioritisation must secure individuals descended from each of the
genetic lineages that comprise each species (Moritz 1994). Previous
attempts to accommodate such into multi-species conservation planning
have partitioned species' ranges into different groups, for example at
habitat discontinuities (e.g. Carvalho \emph{et al.} 2011) as a
pre-processing step to render a new set of ``pseudo-species''. However,
this approach assumes that the biodiversity processes that operate
across a species' range can be readily split into discrete units. Yet
data on biodiversity processes is often continuous and
hyper-dimensional, and often cannot be reduced to a few categories
without significant information loss (Faith \& Walker 1996).

One of the key issues in reserve selection is the lack of a unifying
decision support tool that can accommodate data on biodiversity patterns
and processes in a multi-species context. To begin to fill this gap, we
present the \texttt{rapr R} package. This \texttt{R} package uses novel
formulations of the reserve selection problem to provide decision makers
with the tools to generate prioritisations that preserve both
biodiversity patterns and processes. We aim to define the concepts
behind the problem formulations. Furthermore, we aim to explore the
functionality of the \texttt{rapr R} package by applying one of the
formulations to a set of simulated species and two case studies.

\section{Methods}\label{methods}

\subsection{PROBLEM FORMULATION}\label{problem-formulation}

Biodiversity features are defined as the entities that the
prioritisation is required to preserve (eg. species, ecosystems).
Spatial attributes are defined as the variation across the species'
range that the prioritisation is required to sample. They can be
intrinsic (eg. genetic or trait variation) or extrinsic (eg.
environmental variation) to the feature. These attributes should be
related to the biodiversity processes that the decision maker aims to
preserve.

Each attribute is conceptualised as a space. This space is termed an
``attribute space''. Each planning unit is thought to occupy a single
point inside each space. For example, a decision maker may require a
prioritisation that represents populations along climatic gradients. To
achieve this, the decision maker might use an ``climatic'' attribute
space with dimensions relating to mean annual temperature ($^{\circ}$C)
and precipitation (mm). Any given combination of temperature and
precipitation may be conceived as a point in this environmental space.
By associating planning units with climatic data, they can be mapped
from geographic space to this environmental attribute space.

Demand points are points that also exist in an attribute space. They are
designated by the decision maker to indicate regions of the attribute
space that the decision maker wishes to represent in the prioritisation
(see below for discussion on how demand points can be generated for
real-world datasets). The amount of variation in the attribute space
that a prioritisation secures is a function of the distance between each
demand point and each selected planning unit in the attribute space. The
shorter the distances between the demand points and the planning units;
the better the prioritisation is at securing the variation in the
spatial attribute. To convert these amounts to a proportion--a
meaningful unit for a decision maker--the distances between the selected
planing units and the demand points are scaled by the distances between
the demand points to the centroid of the demand points. In any attribute
space there may exist points that are impossible (eg. mean annual
rainfall -5 mm), or do not occur in the study area (eg. mean annual
temperature 30$^{\circ}$C in Antarctica). Additionally, there may be
some regions that are desirable for some features and undesirable for
others (eg. conditions known to be outside the physiological tolerance
of certain species). Thus a different set of demand points and weights
are used for each attribute space and each feature. By placing demand
points in desirable regions of an attribute space for a given feature,
the decision maker can ensure that prioritisations secure the feature in
planning units with spatial attributes that are desirable for that
feature.

To illustrate these concepts, consider an example conservation planning
scenario example involving an attribute space and demand points. We wish
to develop a prioritisation for a single species that has four
populations in the study area. However, we can only afford to preserve
three populations. We aim preserve the adaptive landscape of the
species, and to achieve this, we will preserve populations inhabiting
different environmental conditions. To describe environmental variation,
we obtain data on the environmental conditions (rainfall (mm) and
temperature ($^{\circ}$C)) where each population is found. These
environmental data are then used to construct a two-dimensional
environmental attribute space. Next, we generated demand points as
equi-distant points between the range of values where the populations
were found. By comparing the distribution of the demand points to the
distribution of the populations in the attribute space, we can identify
a suitable prioritisation (Fig. 1). We can see that preserving both
populations $A$ and $C$, effectively ``doubles-up'' on the same
environmental characteristics, resulting in considerable redundancy in
the solution. Instead, a more representative sample of the
intra-specific variation could be captured by securing populations $A$,
$B$, and $D$. This example demonstrates how the inclusion of
biodiversity processes can guide the reserve selection process.

The formulations used to express the reserve selection problem in the
\texttt{rapr R} package are based on a combination of the
\texttt{Marxan} reserve selection problem and the uncapacitated facility
location problems (Cornu{é}jols \emph{et al.} 1990). Although the
\texttt{rapr R} package provides two novel formulations, for brevity, we
will define the simpler formulation--referred to as the unreliable
formulation--below and define the more complex version--the reliable
formulation--in the Supporting Information S1. These formulations are
named after the unreliable and reliable facility location formulations
from which they are based upon (Cui \emph{et al.} 2010). The key
difference between these two formulations is that the reliable
formulation explicitly considers the probability that planning units are
occupied when calculating the proportion of an attribute space sampled
in a solution. All mathematical terms defined hereafter are described in
Table S1. For convenience, the cardinality of sets will be denoted using
the same symbol used to denote the variable.

Define $F$ to be the set of features one wishes to conserve (indexed by
$f$). Let $J$ be a set of planning units (indexed by $j$). Also, let
$A_j$ denote the area, and $C_j$ denote the cost of preserving planning
unit $j \in J$. To assess the extent to which each feature is secured in
a given prioritisation, let $q_{fj}$ denote the probability of feature
$f$ occupying planning unit $j$. The level of fragmentation associated
with a prioritisation is parametrised as the net exposed boundary
length. Let the shared edges between each planning unit $j \in J$ and
$k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each
$j \in J$ is associated with spatially explicit data that represent
coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a
set of demand points (indexed by $i$) for each feature $f \in F$ and
each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting
for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$
denote the distance between each demand point $i \in I$ and each
planning unit $j \in J$ for each feature $f \in F$ and attribute space
$s \in S$. To describe the inherent variation in the distribution of
demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote
the distance between each demand point $i \in I$ and the centroid of the
demand points. Demand points with greater weight $\lambda_{fsi}$ are
more important, and the optimal solution will be likely to select
planning units close to highly weighted demand points. As a consequence,
the decision maker will need to choose an appropriate weighting for each
demand point.

Targets are used to ensure that prioritisations adequately preserve each
species. Amount-based targets are used to ensure that the total amount
of habitat preserved is sufficient. Let $T_f$ denote the expected amount
of area that needs to be preserved for each feature $f \in F$.
Space-based targets ensure that a sufficient proportion of the
intra-specific variation is secured. Let $\tau_{fs}$ denote the
space-based targets for feature $f \in F$ and attribute space $a \in A$.
For convenience, these both types of targets are expressed as
proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$,
$T_{s}$, and $\tau_{sa}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

The decision variables are the $X_j$ and $Y_{fsij}$ variables.

\begin{align*}
X_j
    &= \begin{cases}
        1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
        0, & \parbox{25em}{otherwise} \\
    \end{cases} \\
%
Y_{fsij} &= \begin{cases}
        1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
        0, & \parbox{25em}{otherwise} \\
    \end{cases} \\
\end{align*}

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is
assigned to a selected planning unit $J$ where $X_j = 1$. The weighted
distance between the demand point and its assigned planning unit
$\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is
represented in a given solution. Generally, demand points are assigned
to the closest selected planning units (unless particularly low
space-based targets are used).

The unreliable formulation (URAP) is a defined as a multi-objective
optimisation problem.

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

The objective function (eqn 3a) determines the utility of a given
prioritisation: a combination of the total cost of a prioritisation and
how fragmented it is. Constraints (eqn 3b) ensure that all the
amount-based targets are met. Constraints (eqn 3c) ensure that all the
space-based targets are met for each feature and each attribute space.
For each feature and attribute space, the total weighted distance
between the demand points and their closest selected planning units is
calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted
distance is then scaled by the inherent variation in the demand points
($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a
proportion conceptually similar to the $R^2$ statistic used in $k$-means
clustering analysis. The constraints ensure that this proportion must be
equal to or greater than the space-based target. Constraints (eqn 3d)
ensure that only one planning unit is assigned to each demand point.
Constraints (eqn 3e) ensure that demand points are only assigned to
selected planning units. Constraints (eqn 3f) ensure that the $X$ and
$Y$ variables are binary.

\subsection{OPTIMISATION}\label{optimisation}

The unreliable formulation is non-linear. However, the non-linear
components can be linearised using existing techniques. The expression
$X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer
et al. (2016). Linearised versions of the problems can be solved using
commercial exact algorithm solvers. The \texttt{rapr} R package provides
functions to express conservation planning data as optimisation problems
using linearised versions of the unreliable and reliable formulations.
These optimisation problems can then be solved to generate
prioritisations using the commercial \texttt{Gurobi} software suite
(\url{http://www.gurobi.com}). Presently, academics can obtain a license
at no cost from the Gurobi website. After installing the \texttt{Gurobi}
software suite, users will need to install the \texttt{Gurobi R}
package.

\section{Examples}\label{examples}

To understand the behaviour of the unreliable problem and showcase its
value, we conducted a simulation study and two case studies. These
studies involved generating solutions using only amount targets to
represent prioritisations generated using conventional methods (eg.
\texttt{Marxan}), and solutions using amount and space targets using the
unreliable formulation. By comparing these solutions, we can guage the
benefits of explicitly including space targets in reserve selection. All
analyses were performed in \texttt{R} (version 3.3.0; {R Core Team}
2016).

\subsection{SIMULATION STUDY}\label{simulation-study}

\subsubsection{Methods}\label{methods-1}

We simulated a hypothetical study area with planning units arranged in
$10 \times 10$ square grid (Fig. 2). We then simulated three species
across this study area. The first species was simulated to represent a
hyper-generalist--occurring in all planning units with equal probability
(Fig. 2a). This species' distribution was based on a uniform
distribution (eqn 8a). The second species was simulated to represent a
species with an idealised distribution (Fig. 2b). It has a core range
area, and is less likely to be found in areas that are distant from the
core area. This species' distribution was simulated using the
probability density function of a single multivariate normal
distribution (eqn 8c). The third species represents a species with two
distinct populations (Fig. 2c). This species' distribution was simulated
to depict a bimodal distribution, based on the combination of the
probability density functions of two multivariate normal distributions
(eqn 8d). The probability that each species inhabited a given planning
unit was calculated using the $X, Y$ coordinates of the planning unit
and eqns 8a--8d. We used a geographic attribute space to showcase the
behavior of the problem. For each species, demand points were generated
by calculating the centroids of each planning units and their weights
were set as the probability that the planning units were occupied.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 8d} \\
%
\end{align*}

We generated four solutions for each species. First to represent
solutions generated using conventional conservation planning methods, we
generated solutions using only 20 \% amount targets. Second to show how
the addition of geographic targets can affect a prioritisation, we
generated solutions using 20 \% amount targets and 75 \% geographic
targets. Third to represent solutions using conventional planning
methods that penalise for fragmentation, we generated solutions using 20
\% amount targets and a boundary length modifier of 1. Fourth to
illustrate the combined affect using geographic targets and penalising
fragmentation, we generated solutions using using 20 \% amount targets
and 75 \% geographic targets and a boundary length modifier of 1.

\subsubsection{Results}\label{results}

The uniform species was simulated to occur with a constant probability
of occupancy across the study area (Fig. 2a). The solution generated
using 20 \% amount-based targets (Fig. 3a) selected 20 planning units
near the southern end of study area. This configuration is an artifact
of the method used to solve this particular instance of the reserve
selection problem. In fact, for this species, all prioritisations
containing 20 planning units are optimal when considering only 20 \%
amount targets. In the absence of criteria to guarantee
representativeness, reserve selection methods may or may not return
solutions that secure a representative sample of the features. In this
particular case, the solution does not secure a representative sample of
the species' range (-23.64 \% sampled).

The addition of a 75 \% geographic target resulted in a solution that
secured a representative sample of the uniform species' range (90.67 \%
sampled; Fig. 3b). Since all planning units have equal cost and an equal
chance of being occupied, this solution has the same number of planning
units as the solution generated using only amount targets (cf.~Fig 3a).
Although the use of amount and space based targets has addressed
adequacy and representativeness objectives (respectively), they have
resulted in a highly fragmented solution.

The addition of a positive boundary length modifier (BLM) parameter
resulted in a well-connected solution that secured an adequate
proportion (20 \% ) and representative sample (75.76 \% sampled) of the
species geographic distribution (Fig. 3d). This solution contains 20
planning units--a few more than the previously discussed solutions--to
ensure that the solution secures a representative proportion of the
species geographic distribution in a configuration that is not highly
fragmented. This result suggests that the combination of space targets
and boundary length modifiers may yield solutions that contain more
planning units, since the solution generated using amount targets and a
boundary length modifier (Fig. 3c) contained the sample number of
planning units as the solution with just the amount targets (Fig. 3a).
Overall, these results show that under the simplest of conditions, the
reliance on just amount targets can cause an under-specified reserve
selection problem that is unlikely to return a suitable solution for
implementation.

The normally distributed species was simulated to contain a single core
area where individuals are most prevalent and marginal areas where
individuals are less likely to occur (Fig. 2b). The solution generated
using 20 \% amount targets contained 10 planning units and concentrated
conservation efforts in the core area (Fig. 3e). Since all planning
units have equal costs and areas, this solution contains the planning
units with the highest probabilities of occupancy. Whilst this solution
satisfies the adequacy objective for a prioritisation in a
cost-effective manner--it fails to fulfill the representative objective
for a prioritisation (59.09 \% distribution sampled).

The solution generated using 20 \% amount targets and 75 \% geographic
targets resulted in a solution that is both adequate and representative
in terms of the uniform species distribution (Fig. 3f). This solution
used 11 planning units to secure 20.02 \% and sample 76.31 \% of the
normal species' distribution. By using amount and geographic targets in
the reserve selection problem, we have obtained a solution that
concentrates conservation effort in the range core and also the range
margin. However, similar to the corresponding solution for the uniform
species (Fig. 3b), this solution is highly fragmented.

By using 20 \% amount targets and 75 \% geographic targets and a
boundary length modifier parameter ($BLM = 1$), we obtained a solution
that fulfills adequacy, representativeness and connectivity objectives
(Fig. 3h). This solution contains 12 planning units. Similar to the
corresponding solution for the uniform species (Fig. 3d), this solution
contains more planning units than any other solutions for this species
(cf.~Figs. 3e--3g). The results for the normally distributed species
suggest that the space targets can result in solutions that secure a
more representative sample of the species'--even for species without
significant structure.

The bimodally distributed species was simulated to represent a species
with a highly structured distribution, and it contains two distinct
populations (Fig. 2c). The solution generated using just 20 \% amount
targets assigned conservation effort to just one of the two populations
(Fig. 3i). While this solution secured an adequate proportion of the
species' distribution (20.18 \% secured), it did not sample a
representative proportion of the range (12 \% sampled). The addition of
boundary length modifiers to just amount targets resulted in a solution
which sampled even less of the species distribution (25.19 \% sampled;
Fig. 3k). However, the addition of geographic space targets resolved
these issues.

The solution generated using 20 \% amount and 75 \% geographic targets
included planning units from both populations (Fig. 3j). Although this
solution required more planning units to fulfill both targets ($n=9$),
this solution secured a representative sample of the bimodally
distributed species (80.95 \% range sampled). Similar to corresponding
solutions generated for both the uniformly (Fig. 3b) and normally
distributed species (Fig. 3f), this solution was also highly fragmented
and we could obtain a more well-connected solution at the expense of
selecting additional planning units ($n=9$; Fig. 3l). The results for
the bimodally distributed species suggest that species with highly
structured populations or strong variation between individuals could
benefit the most from prioritisations that are generated using
space-based targets.

\subsection{CASE STUDY 1}\label{case-study-1}

\subsubsection{Methods}\label{methods-2}

We investigated how space-based targets can be used in a multi-species
planning context to generate a prioritisation that sufficiently
preserves the realised niche for several species. By preserving the
populations in suitable habitats with different with environmental
conditions, conservation planners can preserve the species' adaptive
landscape and foster resilience against environmental change (Moritz
2002). We selected Queensland, Australia as the study area. This region
is ideal for exploring the potential of niche-based targets because it
contains a broad range of different habitats. We obtained data for 19
bioclimatic variables across the region (at $30^{\prime \prime}$
resolution from \url{www.worldclim.org}; Hijmans \emph{et al.} 2005) and
subjected them to a principal components analysis (using ArcMap 10.3.1).
We used the first two principal components (cumulatively explained 99.5
\% of the total variation) to characterise the environmental variation
across the study area (Fig. 4).

We selected four bird species that span a range of different
evolutionary histories, distributions and ecologies: blue-winged
kookaburra (\emph{Dacelo leachii}), brown-backed honeyeater
(\emph{Ramsayornis modestus}), brown falcon (\emph{Falco berigora}), and
pale headed rosella (\emph{Platycercus adscitus}). We then mapped the
extent of occurrence for each species (Fig. 5). To do this, we obtained
observation data from the Atlas of Living Australia across the whole of
Australia (using the \texttt{ALA4R R} package; Raymond \emph{et al.}
2015), spatially thinned the data to omit points within 10 km$^2$ of
each other to ameliorate the effects of sampling bias (using the
developmental version of the \texttt{spThin R} package;
\url{www.github.com/mlammens/spThin}; Aiello-Lammens \emph{et al.}
2015), and fit 85 \% minimum convex polygons (using the
\texttt{adehabitatHR R} package; Calenge 2006). We used this method
because it is trivially reproducible using freely available data.

We generated 500 demand points for each species (Fig. S2). To achieve
this, for each species, we generated random points inside the species
range and at each point extracted the principal component values at that
location. We then fit hyperbox kernels to the distribution of these
points to characterise the realised niche of each species (using a
manually chosen bandwidth of 0.2 and a 0.5 quantile to map the core
parts of the species' niches; implemented in the \texttt{hypervolume R}
package; Blonder \emph{et al.} 2014). We then generated uniformly
distributed points inside the species' kernels and estimated the density
of the training points at the uniformly generated points. These
uniformly distributed points and associated density estimated were used
as demand point coordinates and weights (respectively).

We generated two solutions. The first solution was generated using 20 \%
amount-based targets for all species. The second solution was generated
the same amount-based target with an additional 75 \% niche-based for
each species.

\subsubsection{Results}\label{results-1}

Generally, the solution generated using amount targets preserved a
representative sample of each the four bird species' niches--with one
excpetion. Specifically, this solution sampled over 75 \% of the
realised niche of blue-winged kookaburra (89.54 \%), brown falcon (92.31
\%), and the pale-headed rosella (76.86 \%). Yet it failed to achieve
this for Brown-backed honeyeater(30.93 \%). This result suggests that
prioritisations generated using conventional methods may preserve a
representative sample of most species realised niches (Figure S1).
However, despite this, there may yet be species for which only a small
fraction of their realised niche is preserved. By explictly using space
targets, conservation planners can generate prioritisations that are
guaranteed to capture a representative sample of species' niches.

\subsection{CASE STUDY 2}\label{case-study-2}

\subsubsection{Methods}\label{methods-3}

Here we used space-based targets to generate a prioritisation securing a
representative sample of a species' intra-specific genetic variation. We
used species occurrence and multilocus AFLP data collected by the
international IntraBioDiv project in the European Alps (Meirmans
\emph{et al.} 2011; see Alvarez \emph{et al.} 2009 for further
explanation of data collection methods). Althougthis dataset contains
multiple plant species, we used data for the betony-leaved rampion
(\textit{Phyteuma betonicifolium}), a wide-spread species with
significant genetic structure (Aiello-Lammens \emph{et al.} 2015).
Members of the IntraBioDiv project collected data using a $20^{\prime}$
longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25
km; Fig. 7a). They visited every second grid cell, and if the
betony-leaved rampion was detected in a cell, samples were collected
from three individuals. Samples were genotyped using amplified fragment
length polymorphisms (AFLP; Raymond \emph{et al.} 2015), and used to
construct matrices denoting the presence/absence of polymorphisms at
loci. In total, 131 individuals were genotyped at 138 markers.

We used non-metric multi-dimensional scaling (NMDS; using Gower
distances to accommodate sparsity; Gower 1971; implemented the
\texttt{cluster R} package; Maechler \emph{et al.} 2015) to ordinate the
presence (or absence) of locus-specific alleles wthin individuals into
two continuous variables (implemented in the \texttt{vegan R} package;
Oksanen \emph{et al.} 2015). These continuous variables described the
main axes of genetic variation within the species. We calculated the
average of the values associated with individuals in each grid cell.
These values were used to create a genetic attribute space (Figs. 7c and
7d). To assess spatial auto-correlation, we calculated Moran's I
auto-correlation index for each NMDS axis using inverse great circle
distances based on the grid cells' centroids (using the \texttt{ape R}
package; Paradis \emph{et al.} 2004).

The grid cells were used as planning units. The grid-cell averaged
ordinations were used to describe the typical genetic characteristics of
individuals in the planning unit. Since the number of planning units was
relatively small, we used the same grid-cell averaged ordinations as
demand points. To ensure that the solutions did not prioritise
particularly costly areas, we included opportunity cost data (Fig. 7b).
We obtained population density data (1 km resolution from the Global
Rural-Urban Mapping Project; \texttt{GRUMP V1}; CIESEN, Columbia
University \emph{et al.} 2011) and estimated the total population
density inside each grid cell.

We generated two solutions. The first solution was generated using an 10
\% amount-based target. The second solution was generated using the same
amount-based target and an additional 95 \% genetic-based target.

\subsubsection{Results}\label{results-2}

The binary genetic data was ordinated into a two-dimensional space that
described the main different between individuals (stress = 0.17). These
values were then averaged to the planning unit level to describe the
typical genetic characteristics of individuals in each planning unit
(Figs. 7c--7d). Planning units located near each were found to contain
individuals with similar genetic characteristics (Moran's I: NMDS axis
1, $I = 0.4$, $P < 0.001$; Moran's I: NMDS axis 2, $I = 0.32$,
$P < 0.001$). In terms of the average genetic characteristics of
individuals in the planning units, they tended to cluster into two main
groups, with evidence of within-group structure inside the larger group
(Fig 8c). This analysis supports previous work by Alvarez \emph{et al.}
(2009) who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a
representative portion of the species' genetic variation (15.54 \%
sampled; Figs. 8a and 8c). This solution only sampled planning units
that contained individuals belonging to one of the main two groups. On
the other hand, the solution generated using additional genetic targets
selected planning units belonging to both of the main groups. Note that
since these solutions were generated using opportunity cost data, the
solution generated using only the amount target is essentially the
minimum number of least costly areas needed to fulfill the amount
target. The only difference between the two solutions is a single
planning unit. By swapping this single planning unit, the solution
generated using amount and genetic targets was able to preserve a
representative portion of the species' genetic variation 15.54 for only
a minor increase in cost (99.41 total cost compared to 99.66 total
cost).

\section{Implications and future
directions}\label{implications-and-future-directions}

The \texttt{rapr} R package provides a unified approach to reserve
selection. One of the key advantages of this package is that it is
general enough to incorporate any spatial variable as an attribute
space. This package can accommodate intrinsic or extrinsic variation to
the feature(s). For example, adaptation processes could be secured using
environmental variation or genetic variation among loci under selection.
Ecological processes, such as predator-prey interactions, could be
secured by capturing intra-specific trophic variation. Different
attribute spaces with different targets can be also specified for
different species. As long as the variation can be described using
Euclidean distances, the \texttt{R} package can be used to obtain a
representative sample (note that data may require transformation to
conform to this assumption). This \textbackslash{}texttt\{R\} package
provides decision makers with the tools to generate prioritisations that
secure both biodiversity patterns and the processes that maintain
biodiversity. Additionally, the package contains functionality to
accommodate uncertainty in the distribution of features, and also
identify suitably connected reserves. Both the simulated and case study
species suggest that conservation planning exercises need to explicitly
consider biodiversity processes during the reserve selection process to
ensure they are captured.

The degree to which a prioritisation truly secures a representative
sample of a feature depends on the attribute spaces and distribution of
demand points chosen by the decision maker. Ultimately, the space-based
targets are set as a proportion based on the distribution of the
demand-points. As a consequence, if the decision maker uses an
inappropriate set of spatial variables to construct an attribute space,
or an inappropriate set of demand points, then the optimal solution will
not be an effective prioritisation. We therefore stress that decision
makers must carefully consider which biodiversity processes need to be
reflected in the prioritisation, and which spatial data can be used to
represent these processes. Note that maximising one process can be
detrimental to another. For instance, to maximise the geographic spread
of a prioritisation, reserves need to be further away from each other,
yet to maximise the connectivity of a prioritisation, reserves need to
be closer to each other. To assist in the selection of appropriate
demand points, the \texttt{R} package provides several routines for
generating demand points (see the \texttt{make.DemandPoints} function).
These routines essentially use the distribution of a feature in the
attribute space to define a polygon. Demand points are then generated as
random points within the polygon. A kernel is then fit to the
distribution of the feature in the space (using Blonder \emph{et al.}
2014; Duong 2015), and the demand points are weighted based on the
estimated density of the feature at the demand points.

The formulation requires spatially comprehensive data to map attribute
spaces to planning units. For instance, in both case-studies, all
planning units occupied by the species were associated with
values/coordinates in the attribute spaces. However, most real-world
data sets are patchy--some planning units will be occupied by species
for which attribute space data is not available. To use such patchy data
with the \textbackslash{}texttt\{rapr R\} package, the gaps in the data
must first be filled in as a pre-processing step. Spatially explicit
models could be used to estimate values in locations that are missing
from a data set (eg. krigging or generalised dissimilarity models;
Oliver \& Webster 1990; Ferrier 2002). This approach has been
successfully applied to a range of biological data sets (Thomassen
\emph{et al.} 2010).

To maximise the long-term persistence of biodiversity, decision makers
need to identify prioritisations that preserve existing patterns of
biodiversity and the processes that support them. To achieve this,
conservation planners need a decision support tool that can explicitly
accommodate biodiversity patterns and processes. Here, we developed the
\texttt{rapr R} package to fill this void. By exploring the
functionality of this package using several simulated species, we found
that including space-based targets can radically change a prioritisation
for the simplest of species.

\section{Acknowledgements}\label{acknowledgements}

JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
has an Australian Research Council Future Fellowship. This work was
supported by the Centre of Excellence for Environmental Decisions (CEED)
and the Landscape Ecology and Conservation Group (LEC) at The University
of Queensland.

\section{Data accessibility}\label{data-accessibility}

All data, code, and results are stored in an online repository
(\url{www.github.com/paleo13/rapr-manuscript}) to permit replication and
validation of this study.

\section{References}\label{references}

\bibliography{references}

Aiello-Lammens, M.E., Boria, R.A., Radosavljevic, A., Vilela, B. \&
Anderson, R.P. (2015). spThin: an R package for spatial thinning of
species occurrence records for use in ecological niche models.
\emph{Ecography}, \textbf{38}, 541--545.

Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel, S.,
Sch{ö}nswetter, P., Taberlet, P., Brodbeck, S., Gaudeul, M., Gielly, L.,
K{ü}pfer, P., Mansion, G., Negrini, R., Paun, O., Pellecchia, M., Rioux,
D., Sch{ü}pfer, F., Van Loo, M., Winkler, M., Gugerli, F. \& IntraBioDiv
Consortium. (2009). History or ecology? Substrate type as a major driver
of patial genetic structure in Alpine plants. \emph{Ecology Letters},
\textbf{12}, 632--640.

Ball, I., Possingham, H. \& Watts, M.E. (2009). Marxan and relatives:
software for spatial conservation prioritisation. \emph{Spatial
conservation prioritisation: Quantitative methods \& computational
tools} (eds A. Moilanen, K.A. Wilson \& H. Possingham), pp. 185--189.
Oxford University Press, Oxford, UK.

Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
Solving conservation planning problems with integer linear programming.
\emph{Ecological Modelling}, \textbf{328}, 14--22.

Blonder, B., Lamanna, C., Violle, C. \& Enquist, B.J. (2014). The
\emph{n}-dimensional hypervolume. \emph{Global Ecology and
Biogeography}, \textbf{23}, 595--609.

CIESEN, Columbia University, International Food Policy Research
Institute, The World Bank \& Centro Internacional de Agricultura
Tropical. (2011). Global rural-urban mapping project, Version 1 (GRUMP
v1): Urban extents grid.

Calenge, C. (2006). The package adehabitat for the R software: tool for
the analysis of space and habitat use by animals. \emph{Ecological
Modelling}, \textbf{197}, 1035.

Carvalho, S.B., Brito, J.C., Crespo, E.J. \& Possingham, H.P. (2011).
Incorporating evolutionary processes into conservation planning using
species distribution data: a case study with the western Mediterranean
herpetofauna. \emph{Diversity \& Distributions}, \textbf{17}, 408--421.

Ciarleglio, M., Wesley Barnes, J. \& Sarkar, S. (2009). ConsNet: new
software for the selection of conservation area networks with spatial
and multi-criteria analyses. \emph{Ecography}, \textbf{32}, 205--209.

Cornu{é}jols, G., Nemhauser, G.L. \& Wolsey, L.A. (1990). The
uncapacitated facility location problem. \emph{Discrete Location Theory}
(eds P.B. Mirchandani \& R.L. Francis), pp. 119--171. Wiley, New York.

Crandall, K.A., Bininda-Emonds, O.R.P., Mace, G.M. \& Wayne, R.K.
(2000). Considering evolutionary processes in conservation biology.
\emph{Trends in Ecology \& Evolution}, \textbf{15}, 290--295.

Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
location design under the risk of disruptions. \emph{Operations
Research}, \textbf{58}, 998--1011.

Duong, T. (2015). \emph{ks: Kernel Smoothing, version 1.9.4}.

Faith, D.P. (2003). Environmental diversity (ED) as surrogate
information for species-level biodiversity. \emph{Ecography},
\textbf{26}, 374--379.

Faith, D.P. \& Walker, P.A. (1996). Environmental diversity: on the
best-possible use of surrogate data for assessing the relative
biodiversity of sets of areas. \emph{Biodiversity \& Conservation},
\textbf{5}, 399--415.

Ferrier, S. (2002). Mapping Spatial Pattern in Biodiversity for Regional
Conservation Planning: Where to from Here? \emph{Systematic Biology},
\textbf{51}, 331--363.

Gaston, K.J., Jackson, S.E., Cantu-Salazar, L. \& Cruz-Pinon, G.
(\{2008\}). The Ecological Performance of Protected Areas. \emph{ANNUAL
REVIEW OF ECOLOGY EVOLUTION AND SYSTEMATICS, Annual Review of Ecology
Evolution and Systematics}, \textbf{39}, 93--113.

Gower, J.C. (1971). A general coefficient of similarity and some of its
properties. \emph{Biometrics}, \textbf{27}, 857--871.

Hendry, A.P., Lohmann, L.G., Conti, E., Cracraft, J., Crandall, K.A.,
Faith, D.P., Hauser, C., Joly, C.A., Kogure, K., Larigauderie, A.,
Magallon, S., Moritz, C., Tillier, S., Zardoya, R., Prieur-Richard,
A.H., Walther, B.A., Yahara, T. \& Donoghue, M.J. (2010). Evolutionary
biology in biodiversity science, conservation, and policy: A call to
action. \emph{Evolution}, \textbf{64}, 1517--1528.

Hijmans, R.J., Cameron, S.E., Parra, J.L., Jones, P.G. \& Jarvis, A.
(2005). Very high resolution interpolated climate surfaces for global
land areas. \emph{International Journal of Climatology}, \textbf{25},
1965--1978.

Jones, K.E., Bielby, J., Cardillo, M., Fritz, S.A., O'Dell, J., Orme,
C.D.L., Safi, K., Sechrest, W., Boakes, E.H., Carbone, C., Connolly, C.,
Cutts, M.J., Foster, J.K., Grenyer, R., Habib, M., Plaster, C.A., Price,
S.A., Rigby, E.A., Rist, J., Teacher, A., Bininda-Emonds, O.R.P.,
Gittleman, J.L., Mace, G.M. \& Purvis, A. (2009). PanTHERIA: a
species-level database of life history, ecology, and geography of extant
and recently extinct mammals. \emph{Ecology}, \textbf{90}, 2648--2648.

Klein, C., Wilson, K., Watts, M., Stein, J., Berry, S., Carwardine, J.,
Smith, M.S., Mackey, B. \& Possingham, H. (2009). Incorporating
ecological and evolutionary processes into continental-scale
conservation planning. \emph{Ecological applications}, \textbf{19},
206--217.

Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M. \& Hornik, K.
(2015). \emph{cluster: Cluster Analysis Basics and Extensions}.

Margules, C.R. \& Pressey, R.L. (2000). Systematic conservation
planning. \emph{Nature}, \textbf{405}, 243--253.

McNeely, J.A. (1994). Protected areas for the 21st-century: Working to
provide benefits to society. \emph{Biodiversity and Conservation},
\textbf{3}, 390--405.

Meirmans, P., Goudet, J., IntraBioDiv Consortium \& Gaggiotti, O.
(2011). Ecology and life history affect different aspects of the
population structure of 27 high-alpine plants. \emph{Molecular Ecology},
\textbf{20}, 3144--3155.

Moilanen, A. (2007). Landscape Zonation, benefit functions and
target-based planning: unifying reserve selection strategies.
\emph{Biological Conservation}, \textbf{134}, 571--579.

Moritz, C. (1999). Conservation units and translocations: strategies for
conserving evolutionary processes. \emph{Hereditas}, \textbf{130},
217--228.

Moritz, C. (1994). Defining evolutionarily significant units for
conservation. \emph{Trends in Ecology \& Evolution}, \textbf{9},
373--375.

Moritz, C. (2002). Strategies to protect biological diversity and the
evolutionary processes that sustain it. \emph{Systematic Biology},
\textbf{51}, 238--254.

Oksanen, J., Blanchet, F.G., Kindt, R., Legendre, P., Minchin, P.R.,
O'Hara, R.B., Simpson, G.L., Solymos, P., Stevens, M.H.H. \& Wagner, H.
(2015). \emph{vegan: community ecology package}.

Oliver, M.A. \& Webster, R. (1990). Kriging: a method of interpolation
for geographical information systems. \emph{International Journal of
Geographical Information Systems}, \textbf{4}, 313--332.

Paradis, E., Claude, J. \& Strimmer, K. (2004). APE: analyses of
phylogenetics and evolution in R language. \emph{Bioinformatics},
\textbf{20}, 289--290.

Pyke, C.R. \& Fischer, D.T. (2005). Selection of bioclimatically
representative biological reserve systems under climate change.
\emph{Biological Conservation}, \textbf{121}, 429--441.

{R Core Team}. (2016). R: A language and environment for statistical
computing.

Raymond, B., VanDerWal, J. \& Belbin, L. (2015). \emph{ALA4R: Atlas of
Living Australia (ALA) Data and Resources in R}.

Sanderson, E.W., Segan, D.B. \& Watson, J.E.M. (2015). Global status of
and prospects for protection of terrestrial geophysical diversity.
\emph{Conservation Biology}, \textbf{29}, 649--656.

Thomassen, H.A., Buermann, W., Milá, B., Graham, C.H., Cameron, S.E.,
Schneider, C.J., Pollinger, J.P., Saatchi, S., Wayne, R.K. \& Smith,
T.B. (2010). Modeling environmentally associated morphological and
genetic variation in a rainforest bird, and its application to
conservation prioritization. \emph{Evolutionary Applications},
\textbf{3}, 1--16.

\end{document}