### 2016-11-27 code/rmarkdown/figures.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

---
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
## initialization
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])

# fortify polygons
sim.pus.FPLY <- fortify(sim.pus)
aus.FPLY <- fortify(aus.PLY)
cs1.pus.FPLY <- fortify(cs1.pus)
cs2.grid.FPLY <- fortify(cs2.grid.PLY)
cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)

# store countries
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify

column.width <- 2.75591

```

# Figures

```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Example of an attribute space. This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm). Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found. Points represent demand points. In this space, populations closer to each other have similar environmental conditions."}
## generate data
# population data
populations <- data.frame(
	population=LETTERS[1:4],
	temperature=c(22,29,21.6,34),
	rainfall=c(6,12,5.5,6)
)
# demand point data
make.dp <- function(x, n, pad.percent=0.2) {
	padding<-diff(range(x))*pad.percent
	return(seq(min(x)-padding, max(x)+padding, length.out=n))
}
demand.points <- expand.grid(
	temperature=make.dp(populations$temperature, n=4, pad.percent=0.2),
	rainfall=make.dp(populations$rainfall, n=4, pad.percent=0.2)
)

## plot data
ggplot(aes(x=temperature, y=rainfall), data=populations) +
	geom_point(data=demand.points, fill='gray70') +
	geom_text(aes(label=population), size=6) +
	theme_classic() +
	xlab(expression('Temperature ('*degree*'C)')) +
	ylab('Rainfall (mm)') +
	theme(axis.line.x=element_line(), axis.line.y=element_line())
```

```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.6, fig.width=column.width, fig.cap='Distributions of three simulated species. Squares denote planning units. The color of each planning unit indicates the probability that species inhabit it.'}
# prepare data
sim.pus.occ.FPLY <- ldply(
	seq_len(nrow(sim.species)),
	function(i) {
		curr.DF <- sim.pus.FPLY
		curr.DF$species <- sim.species$name[i]
		curr.DF$Probability <- sim.dps[[i]]@weights[as.numeric(curr.DF$id)]
		return(curr.DF)
	}
)
sim.pus.occ.FPLY$species <- gsub('\n', ' ', as.character(sim.pus.occ.FPLY$species))
sim.pus.occ.FPLY$species <- factor(as.character(sim.pus.occ.FPLY$species), levels=gsub('\n', ' ', sim.species$name)) 
# make letter data.frame
sim.pus.occ.letters.DF <- data.frame(
	x=min(sim.pus.occ.FPLY$long) + 1.5,
	y=max(sim.pus.occ.FPLY$lat) - 1.5,
	species=unique(sim.pus.occ.FPLY$species),
	letter=paste0('(', letters[seq_along(sim.species$name)], ')')
)
# make plot
ggplot() +
	geom_polygon(data=sim.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Probability), color='black') +
	theme_classic() +
	geom_label(data=sim.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=2) +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=7),
    axis.title = element_blank(),
    legend.title = element_text(color='black', size=7),
		legend.text = element_text(color='black', size=7),
		legend.margin = unit(0, 'cm'),
		legend.position='bottom') +
	facet_wrap(~ species, ncol=3)
```

```{r, echo=FALSE, fig.height=3.5, fig.width=2*column.width, fig.cap=paste0('Prioritizations for the simulation study. Each panel shows a prioritization generated for a single species using a set of parameters. Squares denote planning units. Dark green units show planning units selected for protection. The top row of panels (a, d, g, j) shows prioritizations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species. Each column of panels corresponds to a different set of parameters used to generate the prioritisation. The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
# prepare data
sim.pus.sel.FPLY <- ldply(
	seq_len(nrow(sim.spp.DF)),
	function(i) {
		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
		curr.selection <- selections(sim.prioritisations[[as.character(sim.spp.DF$species[i])]][[as.character(sim.spp.DF$prioritisation[i])]])[as.numeric(curr.DF$id)]
		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
		return(curr.DF)
	}
)
sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
# prepare plotting data
sim.letter.DF2 <- sim.spp.DF %>%
  mutate(letter=paste0('(', letters[seq_along(species)],')'),
         long=min(sim.pus.sel.FPLY$long) + 1.6,
         lat=max(sim.pus.sel.FPLY$lat) - 1.5)
# make plot
ggplot() +
	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
	geom_label(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', fill=alpha('white', 0.8), size=5) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank()) +
	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
	facet_grid(species ~ prioritisation)
```

```{r, echo=FALSE, fig.height=1.6, fig.width=column.width, fig.cap='Distribution of environmental variation across Queensland, Australia. Squares denote planning units. Panels show the two main gradients of climatic variation.'}
# prepare data
cs1.pca.FPTS <- ldply(
	seq_len(cs1.params.LST[[MODE]]$n.pc.layers),
	function(i) {
		curr.DF <- cs1.bioclim.RAST[[i]] %>% 
			aggregate(10) %>% 
			rasterToPoints() %>% 
			as.data.frame()
		names(curr.DF)[3] <- 'Score'
		curr.DF$pc <- paste0('PC ',i)
		return(curr.DF)
	}
)
# make plots
p1 <- ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[1]), aes(x=x, y=y, fill=Score)) +
	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
	) +
	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[1])$Score, n=3, min.n=2)) +
	facet_wrap(~ pc)
p2 <- ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[2]), aes(x=x, y=y, fill=Score)) +
	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
		) +
	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[2])$Score, n=2, min.n=2)) +
	facet_wrap(~ pc)
# render plots
grid.arrange(p1, p2, ncol=2)
```

```{r, echo=FALSE, fig.height=2.35, fig.width=column.width, fig.cap='Distribution of the species used in the first case-study. Each panel shows a different species. Squares denote planning units. Planning units occupied by a species are shown in bright blue.'}
# prepare data
cs1.pus.occ.FPLY <- ldply(
	seq_along(cs1.params.LST[[MODE]]$common.names),
	function(i) {
		curr.DF <- cs1.pus.FPLY
		curr.DF$species <- cs1.params.LST[[MODE]]$common.names[i]
		curr.DF$Occupancy <- 'absent'
		curr.DF$Occupancy[which(curr.DF$id %in% filter(cs1.ru@data@pu.species.probabilities, species==i)$pu)] <- 'present'
		return(curr.DF)
	}
)
cs1.pus.occ.FPLY$species <- factor(as.character(cs1.pus.occ.FPLY$species), levels=cs1.params.LST[[MODE]]$common.names)
# make letter data.frame
cs1.pus.occ.letters.DF <- data.frame(
	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.06)),
	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.90)),
	species=unique(cs1.pus.occ.FPLY$species),
	letter=paste0('(', letters[seq_along(unique(cs1.pus.occ.FPLY$species))], ')')
)
# make plots
ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black', size=0.05) +
	geom_label(data=cs1.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		axis.title=element_blank(), legend.position='none'
	) +
	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
	facet_wrap(~ species, ncol=2)
```

```{r, echo=FALSE, fig.height=1.175, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study. Squares represent planning units. Dark green planning units were selected for protection. Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets. Panel (b) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
# prepare data
cs1.pus.sel.FPLY <- ldply(
	names(cs1.prioritisations),
	function(x) {
		curr.DF <- cs1.pus.FPLY
		curr.DF$Prioritisation <- x
		curr.DF$Solution <- 'discarded'
		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
		return(curr.DF)
	}
)
cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
# make letter data.frame
cs1.pus.sel.letters.DF <- data.frame(
	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.06)),
	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.90)),
	Prioritisation=names(cs1.prioritisations),
	letter=paste0('(', letters[1:2], ')')
)
# make plots
ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black', size=0.1) +
	geom_label(data=cs1.pus.sel.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		axis.title=element_blank(), legend.position='none'
	) +
	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
	facet_wrap(~ Prioritisation, ncol=2)
```

```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Data used for the second case-study. Squares denote planning units. Panel (a) shows all grid cells surveyed by the IntraBioDiv project. Grid cells occupied by ',tolower(cs2.params.LST[[MODE]]$common.name),' are shown in bright blue. The subsequent panels contain only the grid cells that were occupied by the species. Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density). Panels (c--d) show the spatial distribution of the ordinated genetic data. These values describe the typical genetic characteristics of individuals in each planning unit. Planning units with similar values/colors contain individuals with similar loci polymorphisms.')}
# prepare full grid data
cs2.spp.grid.FPLY <- cs2.grid.FPLY
cs2.spp.grid.FPLY$Title <- 'Occupancy'
cs2.spp.grid.FPLY$Occupancy <- 'absent'
cs2.spp.grid.FPLY$Occupancy[which(cs2.grid.PLY@data[[cs2.params.LST[[MODE]]$species.name]][as.numeric(cs2.spp.grid.FPLY$id)]==1)] <- 'present'

cs2.spp.grid.letters.DF <- data.frame(
	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.06)),
	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.90)),
	letter=('(a)')
)

# prepare subset grid data
cs2.spp.grid.sub.FPLY <- cs2.grid.sub.FPLY
cs2.spp.grid.sub.FPLY$nmds1 <- NA
cs2.spp.grid.sub.FPLY$nmds2 <- NA
cs2.nmds1 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d1')]]
cs2.nmds2 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d2')]]

curr.pos <- match(as.numeric(cs2.spp.grid.sub.FPLY$id), cs2.grid.sub.DF$id)
cs2.spp.grid.sub.FPLY$nmds1 <- cs2.nmds1[curr.pos]
cs2.spp.grid.sub.FPLY$nmds2 <- cs2.nmds2[curr.pos]

cs2.spp.grid.sub.FPLY$Cost <- costs.MTX[match(cs2.grid.sub.PLY$id[as.numeric(cs2.spp.grid.sub.FPLY$id)], costs.MTX[,1]),2]

cs2.spp.grid.sub.FPLY$Title1 <- 'Opportunity cost'
cs2.spp.grid.sub.FPLY$Title2 <- 'Genetic NMDS 1'
cs2.spp.grid.sub.FPLY$Title3 <- 'Genetic NMDS 2'

cs2.spp.grid.sub.letters.DF <- data.frame(
	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.90)),
	letter=c('(b)','(c)','(d)')
)

# plot species data
p1 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_polygon(data=cs2.spp.grid.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), alpha=0.8, color='grey10') +
	geom_label(data=cs2.spp.grid.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank(), legend.position='none'
	) +
	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
	coord_cartesian(xlim=buffered.range(cs2.grid.FPLY$long, 0), ylim=buffered.range(cs2.grid.FPLY$lat, 0)) +
	facet_wrap(~ Title)
# plot cost data
p2 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Cost), alpha=0.8, color='grey10') +
	geom_label(data=cs2.spp.grid.sub.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
		legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
		legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
	) +
	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
	scale_fill_continuous(name=element_blank(), breaks=pretty(cs2.spp.grid.sub.FPLY$Cost, n=3, min.n=2)) +
	facet_wrap(~ Title1)
# plot nmds axis 1
p3 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds1), alpha=0.8, color='grey10') +
	geom_label(data=cs2.spp.grid.sub.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
    axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
    legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
    legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
	) + 
	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred', breaks=c(0, 0.3)) +
	facet_wrap(~ Title2)
# plot nmds axis 2
p4 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds2), alpha=0.8, color='grey10') +
	geom_label(data=cs2.spp.grid.sub.letters.DF[3,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
    axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
    legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
    legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
	) + 
	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred', breaks=pretty(cs2.spp.grid.sub.FPLY$nmds2, n=2, min.n=2)) +
	facet_wrap(~ Title3)
# render plots
grid.arrange(p1, p2, p3, p4, ncol=2)
```

```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Prioritizations for the second case study. Panels (a--b) show prioritisations generated using different parameters. Squares denote planning units shown in geographic space. Dark green planning units were selected for protection. Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets. Panel (b) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs2.params.LST[[MODE]]$genetic.target*100,' % genetic targets. Panels (c--d) show the solutions in the genetic space. Each point corresponds to a planning unit. The coordinates of the points represent the typical genetic characeristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data). Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.')}
# prepare spatial data
cs2.sel.grid.sub.FPLY <- cs2.grid.sub.FPLY
cs2.sel.grid.sub.FPLY$Title1 <-'Amount target'
cs2.sel.grid.sub.FPLY$Title2 <- 'Amount & genetic target'

cs2.sel.grid.sub.FPLY$Amount.selection <- 'discarded'
cs2.sel.grid.sub.FPLY$Amount.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'

cs2.sel.grid.sub.FPLY$Amount.genetic.selection <- 'discarded'
cs2.sel.grid.sub.FPLY$Amount.genetic.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'

cs2.sel.grid.sub.letters.DF <- data.frame(
	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.90)),
	letter=c('(a)','(b)')
)

# prepare genetic space plots
cs2.sel.space.DF <- data.frame(cs2.genetic.AS@spaces[[1]]@planning.unit.points@coords)
names(cs2.sel.space.DF) <- c('d1', 'd2')
cs2.sel.space.DF$id <- cs2.genetic.AS@spaces[[1]]@planning.unit.points@ids

cs2.sel.space.DF$Amount.selection <- 'discarded'
cs2.sel.space.DF$Amount.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'

cs2.sel.space.DF$Amount.genetic.selection <- 'discarded'
cs2.sel.space.DF$Amount.genetic.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'

cs2.sel.space.letters.DF <- data.frame(
	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.85),
	y=min(cs2.sel.space.DF$d2) + (diff(range(cs2.sel.space.DF$d2))*0.95),
	letter=c('(c)', '(d)')
)

# make plots
p1 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.selection), color='black') +
	geom_label(data=cs2.sel.grid.sub.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
		axis.title=element_blank(), legend.position='none'
	) +
	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
	facet_wrap(~ Title1)
p2 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.genetic.selection), color='black') +
	geom_label(data=cs2.sel.grid.sub.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'), legend.position='none'
	) +
	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
	facet_wrap(~ Title2)
p3 <- ggplot() +
	geom_point(data=arrange(cs2.sel.space.DF, Amount.selection), aes(x=d1, y=d2, fill=Amount.selection), pch=21, size=3, color='black') +
	geom_label(data=cs2.sel.space.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() + xlab('NMDS 1') + ylab('NMDS 2') +
	scale_x_continuous(expand=c(0.08, 0)) + scale_y_continuous(expand=c(0.08, 0)) + 
	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line(),
    axis.text=element_text(size=8), axis.title=element_text(size=8), plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
p4 <- ggplot() +
	geom_point(data=arrange(cs2.sel.space.DF, Amount.genetic.selection), aes(x=d1, y=d2, fill=Amount.genetic.selection), pch=21, size=3, color='black') +
	geom_label(data=cs2.sel.space.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() + xlab('NMDS 1') + ylab('NMDS 2') +
	scale_x_continuous(expand=c(0.08, 0)) + scale_y_continuous(expand=c(0.08, 0)) + 
  theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line(),
    axis.text=element_text(size=8), axis.title=element_text(size=8), plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
# render plots
grid.arrange(p1,p2,p3,p4,ncol=2)
```