### 2017-04-06 code/rmarkdown/figures.Rmd
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

---
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r, include = FALSE}
# set wd
knitr::opts_knit$set(root.dir = normalizePath("../.."))
```

```{r, include = FALSE}
## initialization
# load .rda file
try(session::restore.session("data/final/results.rda"))

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(raster::nlayers(cs1_bioclim_raster)))
  cs1_bioclim_raster@layers[[i]]@file@name  <- paste0(getwd(),
    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
             "raptr-manuscript")[[1]][[2]])

# remove small islands from australia basemap
aus_sub_polygons <- sp::disaggregate(aus_polygons)
aus_sub_polygons <- aus_sub_polygons %>%
                    sp::SpatialPolygonsDataFrame(
                      data = data.frame(size = rgeos::gArea(aus_sub_polygons,
                                                            byid = TRUE)))
aus_sub_polygons <- aus_sub_polygons[aus_sub_polygons$size >=
  quantile(aus_sub_polygons$size, 0.9, names=FALSE),]

# fortify polygons
sim_pus_fpolygons <- ggplot2::fortify(sim_pus)
aus_fpolygons <- ggplot2::fortify(aus_sub_polygons)
cs1_pus_fpolygons <- ggplot2::fortify(cs1_pus)
cs2_grid_fpolygons <- ggplot2::fortify(cs2_grid_polygons)
cs2_grid_sub_fpolygons <- ggplot2::fortify(cs2_grid_sub_polygons)

# store countries
data(countriesHigh, package = "rworldxtra")
countries_fpolygons <- countriesHigh[
  countriesHigh$ADMIN %in% c(
    "Italy", "Switzerland", "France", "Austria",
    "Germany", "Slovenia", "Croatia", "Hungary",
    "Monaco", "Germany", "Slovakia", "Czech Republic"
  ),] %>% sp_fortify

column_width <- 2.75591

```

# Figures

```{r, echo = FALSE, message = FALSE, fig.height = 2, fig.width = 2, fig.cap = "Attribute space example. This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm). Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found. Points represent demand points. In this space, populations close to each other inhabit similar environmental conditions."}
## generate data
# population data
populations <- data.frame(population = LETTERS[1:4],
                          temperature = c(35, 34, 31, 21.6),
                          rainfall = c(4.5, 5.3, 9.5, 5.5))

# demand point data
make_dp <- function(x, n, pad.percent = 0.2) {
  padding <- diff(range(x)) * pad.percent
  return(seq(min(x) - padding, max(x) + padding, length.out = n))
}
demand.points <- expand.grid(
  temperature = make_dp(populations$temperature, n = 4, pad.percent = 0.2),
  rainfall = make_dp(populations$rainfall, n = 4, pad.percent = 0.2))

## plot data
ggplot2::ggplot(ggplot2::aes(x = rainfall, y = temperature),
                data = populations) +
  ggplot2::geom_point(data = demand.points, fill = "gray70") +
  ggplot2::geom_text(ggplot2::aes(label = population), size = 6) +
  ggplot2::theme_classic() +
  ggplot2::ylab(expression("Temperature ("*degree*"C)")) +
  ggplot2::xlab("Rainfall (mm)") +
  ggplot2::theme(axis.line.x = ggplot2::element_line(),
                 axis.line.y = ggplot2::element_line())
```

```{r, echo = FALSE, messages = FALSE, results = "hide", fig.height = 1.6, fig.width = column_width, fig.cap = "Distributions of three simulated species. Squares denote planning units. Colors indicate probability of occupancy."}
# prepare data
sim_pus_occ_fpolygons <- plyr::ldply(
  seq_len(nrow(sim_species)),
  function(i) {
    curr_data <- sim_pus_fpolygons
    curr_data$species <- sim_species$name[i]
    curr_data$Probability <- sim_demand_points[[i]]@weights[
      as.numeric(curr_data$id)]
    return(curr_data)
  }
)
sim_pus_occ_fpolygons$species <- gsub("\n", " ",
  as.character(sim_pus_occ_fpolygons$species))
sim_pus_occ_fpolygons$species <- factor(as.character(
  sim_pus_occ_fpolygons$species), levels = gsub("\n", " ", sim_species$name))
# make letter data.frame
sim_pus_occ_letters <- data.frame(
  x = min(sim_pus_occ_fpolygons$long) + 1.5,
  y = max(sim_pus_occ_fpolygons$lat) - 1.5,
  species = unique(sim_pus_occ_fpolygons$species),
  letter=paste0("(", letters[seq_along(sim_species$name)], ")"))
# make plot
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = sim_pus_occ_fpolygons,
                        ggplot2::aes(x = long, y = lat,  group = group,
                                     fill = Probability),
                        color="black") +
  ggplot2::theme_classic() +
  ggplot2::geom_label(data = sim_pus_occ_letters,
                     ggplot2::aes(label = letter, x = x, y =y ), color="black",
                     fill = scales::alpha("white", 0.8), size = 2) +
  ggplot2::theme(axis.ticks = ggplot2::element_blank(),
                axis.text = ggplot2::element_blank(),
                axis.line = ggplot2::element_blank(),
                strip.background = ggplot2::element_rect(fill = "grey20"),
                strip.text = ggplot2::element_text(color = "white", size = 7),
                axis.title = ggplot2::element_blank(),
                legend.text = ggplot2::element_text(color = "black", size = 7),
                legend.margin = ggplot2::unit(0, "cm"),
                legend.position = "bottom") +
  ggplot2::facet_wrap(~ species, ncol = 3)
```

```{r, echo = FALSE, fig.height = 3.5, fig.width = 2 * column_width, fig.cap = paste0("Prioritizations for the simulation study. Each panel shows a prioritization generated for a single species using a set of parameters. Squares denote planning units. Dark green planning units were selected for protection. Each row of panels show prioritizations generated for a different species. Each column of panels corresponds to a different set of parameters used to generate the prioritization.")}
# prepare data
sim_pus_sel_fpolygons <- plyr::ldply(seq_len(nrow(sim_spp_results)), function(i) {
    curr_data <- sim_pus_fpolygons %>%
                 dplyr::mutate(species = sim_spp_results$species[i],
                               prioritisation =
                                sim_spp_results$prioritisation[i])
    curr_selection <- sim_prioritisations %>%
                      `[[`(as.character(sim_spp_results$species[i])) %>%
                      `[[`(as.character(sim_spp_results$prioritisation[i])) %>%
                      raptr::selections() %>%
                      `[`(as.numeric(curr_data$id))
    curr_data$Selection <-  c("discarded","selected")[curr_selection+1]
    return(curr_data)
})
sim_pus_sel_fpolygons$prioritisation <- factor(
  as.character(sim_pus_sel_fpolygons$prioritisation),
  levels = c("Amount\ntargets", "Amount &\nspace targets",
             "Amount target\nand BLM", "Amount & space\ntargets and BLM"))
# prepare plotting data
sim_letter_data2 <- sim_spp_results %>%
  dplyr::mutate(letter = paste0("(", letters[seq_along(species)],")"),
                long = min(sim_pus_sel_fpolygons$long) + 1.6,
                lat = max(sim_pus_sel_fpolygons$lat) - 1.5)
# make plot
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = sim_pus_sel_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Selection),
                         color="black") +
  ggplot2::geom_label(data = sim_letter_data2,
                      ggplot2::aes(label = letter, x = long, y = lat),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size=5) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    axis.line.x = ggplot2::element_blank(),
    axis.line.y = ggplot2::element_blank(),
    plot.background = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    panel.background = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    legend.position="none",
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank()) +
  ggplot2::scale_fill_manual(name="Solution",
                             values = c("discarded"="#f7fcfd",
                                        "selected"="#00441b")) +
  ggplot2::facet_grid(species ~ prioritisation)
```

```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia. Polygons denote planning units."}
# prepare data
cs1_pca_fpoints <- plyr::ldply(seq_len(cs1_parameters$number_pc_layers),
  function(i) {
    curr_data <- cs1_bioclim_raster[[i]] %>%
                     raster::aggregate(10) %>%
                     raster::rasterToPoints() %>%
                     as.data.frame()
    names(curr_data)[3] <- "Score"
    curr_data$pc <- paste0("PC ",i)
    return(curr_data)
})
# make plots
p1 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[1]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = NA, color = "black", size = 0.05) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color = "transparent",
                                              fill = "transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[1])$Score, n = 3,
          min.n = 2)) +
  ggplot2::facet_wrap(~ pc)
p2 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill="grey85", color="grey70") +
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[2]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = NA, color = "black", size = 0.05) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color="transparent",
                                              fill="transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[2])$Score, n = 2,
           min.n = 2)) +
  ggplot2::facet_wrap(~ pc)
# render plots
gridExtra::grid.arrange(p1, p2, ncol = 2)
```

```{r, echo = FALSE, fig.height = 3.8, fig.width = column_width, fig.cap = "Distribution of the species used in the first case study. Polygons denote planning units. Planning units occupied by a given species are shown in light blue."}
# prepare data
cs1_pus_occ_fpolygons <- plyr::ldply(
  seq_along(cs1_parameters$common_names),
  function(i) {
    curr_data <- cs1_pus_fpolygons
    curr_data$species <- cs1_parameters$common_names[i]
    curr_data$Occupancy <- "absent"
    curr_pus <- cs1_ru@data@pu.species.probabilities %>%
                dplyr::filter(species == i) %>%
                `[[`("pu")
    curr_data$Occupancy[which(curr_data$id %in% curr_pus)] <- "present"
    return(curr_data)
})

cs1_pus_occ_fpolygons$species <- factor(
    as.character(cs1_pus_occ_fpolygons$species),
    levels = cs1_parameters$common_names)
# make letter data.frame
cs1_pus_occ_letters_data <- data.frame(
  x = unname(cs1_pus@bbox[1,1] + (diff(cs1_pus@bbox[1, ]) * 0.08)),
  y = unname(cs1_pus@bbox[2,1] + (diff(cs1_pus@bbox[2, ]) * 0.95)),
  species=unique(cs1_pus_occ_fpolygons$species),
  letter=paste0("(", letters[seq_along(unique(cs1_pus_occ_fpolygons$species))],
                ")")
)
# make plots
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70", size = 0.05) +
  ggplot2::geom_polygon(data = cs1_pus_occ_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Occupancy),
                        color = "black", size = 0.05) +
  ggplot2::geom_label(data = cs1_pus_occ_letters_data,
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8), size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white", size = 7),
    axis.title = ggplot2::element_blank(),
    legend.position = "none") +
  ggplot2::scale_fill_manual(name = "Occupancy",
                             values = c("absent"="#132B43",
                                        "present"="#56B1F7")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::facet_wrap(~ species, ncol = 2)
```

```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = paste0("Prioritizations for the first case study. Polygons denote planning units. Dark green planning units were selected for protection. Panel (a) shows the solution generated when using ",cs1_parameters$amount_target*100," % amount targets. Panel (b) shows the solution when using ", cs1_parameters$amount_target*100, " % amount targets and ", cs1_parameters$space_target*100, " % niche targets.")}
# prepare data
cs1_pus_sel_fpolygons <- plyr::ldply(
  names(cs1_prioritisations),
  function(x) {
    curr_data <- cs1_pus_fpolygons
    curr_data$Prioritisation <- x
    curr_data$Solution <- "discarded"
    curr_pus <- which(as.logical(raptr::selections(cs1_prioritisations[[x]])))
    curr_data$Solution[curr_data$id %in% curr_pus] <- "selected"
    return(curr_data)
  }
)
cs1_pus_sel_fpolygons$Prioritisation <- factor(as.character(
  cs1_pus_sel_fpolygons$Prioritisation),
  levels = names(cs1_prioritisations))
# make letter data.frame
cs1_pus_sel_letters_data <- data.frame(
  x = unname(cs1_pus@bbox[1,1] + (diff(cs1_pus@bbox[1, ]) * 0.08)),
  y = unname(cs1_pus@bbox[2,1] + (diff(cs1_pus@bbox[2, ]) * 0.95)),
  Prioritisation = names(cs1_prioritisations),
  letter = paste0("(", letters[1:2], ")"))
# make plots
ggplot2::ggplot() +
  ggplot2::geom_polygon(data=aus_fpolygons,
                        ggplot2::aes(x=long, y=lat, group=group),
                        fill = "grey85", color = "grey70", size = 0.05) +
  ggplot2::geom_polygon(data = cs1_pus_sel_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                    fill = Solution),
                        color = "black", size = 0.05) +
  ggplot2::geom_label(data = cs1_pus_sel_letters_data,
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                                                             size=3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white", size = 8),
    axis.title = ggplot2::element_blank(),
    legend.position = "none") +
  ggplot2::scale_fill_manual(name = "Solution",
                             values = c("discarded"="#f7fcfd",
                                        "selected"="#00441b")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::facet_wrap(~ Prioritisation, ncol = 2)
```

```{r, echo = FALSE, fig.height = 2.2, fig.width = column_width, fig.cap = paste0("Data used for the second case study. Squares denote planning units. Panel (a) shows all grid cells surveyed by the IntraBioDiv project. Grid cells occupied by the ",tolower(cs2_parameters$common_name)," are shown in bright blue. The subsequent panels contain only show occupied grid cells. Panel (b) shows the acquisition cost of each planning unit (estimated as the total human population density). Panels (c--d) show the spatial distribution of the ordinations describing genetic variation. These values describe the typical genetic characteristics of individuals in each planning unit. Planning units with similar values/colors contain individuals with similar loci polymorphisms.")}
# prepare full grid data
cs2_spp_grid_fpolygons <- cs2_grid_fpolygons
cs2_spp_grid_fpolygons$Title <- "Occupancy"
cs2_spp_grid_fpolygons$Occupancy <- "absent"
status <- cs2_grid_polygons@data[[cs2_parameters$species_name]][as.numeric(
        cs2_spp_grid_fpolygons$id)]
cs2_spp_grid_fpolygons$Occupancy[which(status==1)] <- "present"

cs2_spp_grid_letters_data <- data.frame(
  x = unname(cs2_grid_polygons@bbox[1, 1] +
      (diff(cs2_grid_polygons@bbox[1, ]) * 0.07)),
  y = unname(cs2_grid_polygons@bbox[2, 1] +
      (diff(cs2_grid_polygons@bbox[2, ]) * 0.875)),
  letter = ("(a)"))

# prepare subset grid data
cs2_spp_grid_sub_fpolygons <- cs2_grid_sub_fpolygons
cs2_spp_grid_sub_fpolygons$nmds1 <- NA
cs2_spp_grid_sub_fpolygons$nmds2 <- NA
cs2.nmds1 <- cs2_grid_sub_data[[paste0(cs2_parameters$species_name,
                                       "_genetic_d1")]]
cs2.nmds2 <- cs2_grid_sub_data[[paste0(cs2_parameters$species_name,
                                       "_genetic_d2")]]

curr_pos <- match(as.numeric(cs2_spp_grid_sub_fpolygons$id),
                  cs2_grid_sub_data$id)
cs2_spp_grid_sub_fpolygons$nmds1 <- cs2.nmds1[curr_pos]
cs2_spp_grid_sub_fpolygons$nmds2 <- cs2.nmds2[curr_pos]

ids <- cs2_grid_sub_polygons$id[as.numeric(cs2_spp_grid_sub_fpolygons$id)]
cs2_spp_grid_sub_fpolygons$Cost <- costs_matrix[match(ids,
                                                      costs_matrix[, 1]), 2]

cs2_spp_grid_sub_fpolygons$Title1 <- "Opportunity cost"
cs2_spp_grid_sub_fpolygons$Title2 <- "Genetic NMDS 1"
cs2_spp_grid_sub_fpolygons$Title3 <- "Genetic NMDS 2"

cs2_spp_grid_sub_letters_data <- data.frame(
  x = unname(cs2_grid_sub_polygons@bbox[1, 1] +
      (diff(cs2_grid_sub_polygons@bbox[1, ]) * 0.06)),
  y = unname(cs2_grid_sub_polygons@bbox[2, 1] +
      (diff(cs2_grid_sub_polygons@bbox[2, ]) * 0.89)),
  letter = c("(b)", "(c)", "(d)"))

# plot species data
p1 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = countries_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_polygon(data = cs2_spp_grid_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Occupancy),
                        alpha = 0.8, color = "grey10") +
  ggplot2::geom_label(data = cs2_spp_grid_letters_data,
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.position = "none") +
  ggplot2::scale_fill_manual(name = "Occupancy",
                             values = c("absent"="#132B43",
                                        "present"="#56B1F7")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs2_grid_fpolygons$long, 0),
                           ylim = buffered_range(cs2_grid_fpolygons$lat, 0)) +
  ggplot2::facet_wrap(~ Title)
# plot cost data
p2 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = countries_fpolygons,
                        ggplot2::aes(x =long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Cost),
                        alpha = 0.8, color = "grey10") +
  ggplot2::geom_label(data = cs2_spp_grid_sub_letters_data[1, ],
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill="grey20"),
    strip.text = ggplot2::element_text(color="white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color = "black", fill = "white"),
    legend.position = c(0.25, 0.58),
    legend.key.height = ggplot2::unit(2.5, "mm"),
    legend.key.width = ggplot2::unit(3.25, "mm"),
    legend.text = ggplot2::element_text(size = 6.0),
    legend.title = ggplot2::element_blank(),
    legend.direction = "horizontal",
    legend.justification = c(0,1),
    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
                                                                  "mm"))) +
  ggplot2::coord_cartesian(
    xlim = buffered_range(cs2_grid_sub_fpolygons$long, 0),
    ylim = buffered_range(cs2_grid_sub_fpolygons$lat, 0)) +
    ggplot2::scale_fill_continuous(name = ggplot2::element_blank(),
      breaks = pretty(cs2_spp_grid_sub_fpolygons$Cost, n = 3, min.n = 2)) +
  ggplot2::facet_wrap(~ Title1)
# plot nmds axis 1
p3 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = countries_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = nmds1),
                         alpha=0.8, color="grey10") +
  ggplot2::geom_label(data = cs2_spp_grid_sub_letters_data[2, ],
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill="grey20"),
    strip.text = ggplot2::element_text(color="white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color="black", fill="white"),
    legend.position = c(0.25, 0.58),
    legend.key.height = ggplot2::unit(2.5, "mm"),
    legend.key.width = ggplot2::unit(3.25, "mm"),
    legend.text = ggplot2::element_text(size=6.0),
    legend.title = ggplot2::element_blank(),
    legend.direction = "horizontal",
    legend.justification = c(0,1),
    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
                                                                  "mm"))) +
  ggplot2::coord_cartesian(
      xlim = buffered_range(cs2_grid_sub_fpolygons$long, 0),
      ylim = buffered_range(cs2_grid_sub_fpolygons$lat, 0)) +
  ggplot2::scale_fill_gradient2(name = ggplot2::element_blank(),
                                low = "darkblue", mid = "white",
                                high = "darkred", breaks = c(0, 0.3)) +
  ggplot2::facet_wrap(~ Title2)
# plot nmds axis 2
p4 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = countries_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill="grey85", color="grey70") +
  ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = nmds2),
                         alpha=0.8, color="grey10") +
  ggplot2::geom_label(data = cs2_spp_grid_sub_letters_data[3,],
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill="grey20"),
    strip.text = ggplot2::element_text(color="white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color="black", fill="white"),
    legend.position = c(0.25, 0.58),
    legend.key.height = ggplot2::unit(2.5, "mm"),
    legend.key.width = ggplot2::unit(3.25, "mm"),
    legend.text = ggplot2::element_text(size=6.0),
    legend.title = ggplot2::element_blank(),
    legend.direction = "horizontal",
    legend.justification = c(0,1),
    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
                                                                   "mm"))) +
  ggplot2::coord_cartesian(
    xlim = buffered_range(cs2_grid_sub_fpolygons$long, 0),
    ylim = buffered_range(cs2_grid_sub_fpolygons$lat, 0)) +
  ggplot2::scale_fill_gradient2(name = ggplot2::element_blank(),
                                low = "darkblue", mid = "white",
                                high = "darkred",
                                breaks = pretty(
                                  cs2_spp_grid_sub_fpolygons$nmds2, n = 2,
                                  min.n = 2)) +
  ggplot2::facet_wrap(~ Title3)
# render plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

```{r, echo = FALSE, fig.height = 2.5, fig.width = column_width, fig.cap = paste0("Prioritizations for the second case study. Panels (a--b) show prioritizations generated using different parameters. Polygons denote planning units. Dark green planning units were selected for protection. Panel (a) shows the planning units selected when using ",cs2_parameters$amount_target * 100," % amount targets. Panel (b) shows the planning units selected when using ",cs2_parameters$amount.target*100," % amount targets and ",cs2_parameters$genetic_target * 100," % genetic targets. Panels (c--d) show the solutions in the genetic space. Each point corresponds to a planning unit. The coordinates of the points represent the typical genetic characteristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data). Planning units associated with points that are closer together contain individuals with more similar genetic characteristics than planning units that are further apart.")}
# prepare spatial data
cs2_sel_grid_sub_fpolygons <- cs2_grid_sub_fpolygons
cs2_sel_grid_sub_fpolygons$Title1 <-"Amount target"
cs2_sel_grid_sub_fpolygons$Title2 <- "Amount & genetic target"

cs2_sel_grid_sub_fpolygons$Amount_selection <- "discarded"
sel_1 <- as.logical(raptr::selections(cs2_prioritisations[[1]]))
rows_1 <- cs2_sel_grid_sub_fpolygons$id %in% which(sel_1)
cs2_sel_grid_sub_fpolygons$Amount_selection[rows_1] <- "selected"

cs2_sel_grid_sub_fpolygons$Amount_genetic_selection <- "discarded"
sel_2 <- as.logical(raptr::selections(cs2_prioritisations[[2]]))
rows_2 <- cs2_sel_grid_sub_fpolygons$id %in% which(sel_2)
cs2_sel_grid_sub_fpolygons$Amount_genetic_selection[rows_2] <- "selected"

cs2_sel_grid_sub_letters_data <- data.frame(
  x = unname(cs2_grid_sub_polygons@bbox[1, 1] +
      (diff(cs2_grid_sub_polygons@bbox[1, ]) * 0.06)),
  y = unname(cs2_grid_sub_polygons@bbox[2,1] +
      (diff(cs2_grid_sub_polygons@bbox[2, ]) * 0.9)),
  letter = c("(a)", "(b)"))

# prepare genetic space plots
cs2_sel_space_data <- cs2_genetic_attribute_spaces %>%
                      slot("spaces") %>%
                      `[[`(1) %>%
                      slot("planning.unit.points") %>%
                      slot("coords") %>%
                      data.frame() %>%
                      magrittr::set_names(c("d1", "d2")) %>%
                      dplyr::mutate(id = cs2_genetic_attribute_spaces %>%
                                         slot("spaces") %>%
                                         `[[`(1) %>%
                                         slot("planning.unit.points") %>%
                                         slot("ids"))

cs2_sel_space_data$Amount_selection <- "discarded"
sel_1 <- as.logical(raptr::selections(cs2_prioritisations[[1]]))
rows_1 <- cs2_sel_space_data$id %in% which(sel_1)
cs2_sel_space_data$Amount_selection[rows_1] <- "selected"

cs2_sel_space_data$Amount_genetic_selection <- "discarded"
sel_2 <- as.logical(raptr::selections(cs2_prioritisations[[2]]))
rows_2 <- cs2_sel_space_data$id %in% which(sel_2)
cs2_sel_space_data$Amount_genetic_selection[rows_2] <- "selected"

cs2_sel_space_letters_data <- data.frame(
  x = min(cs2_sel_space_data$d1) +
      (diff(range(cs2_sel_space_data$d1)) * 0.85),
  y = min(cs2_sel_space_data$d2) +
      (diff(range(cs2_sel_space_data$d2)) * 0.95),
  letter = c("(c)", "(d)"))

# make plots
p1 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = countries_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_polygon(data = cs2_sel_grid_sub_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Amount_selection), color="black") +
  ggplot2::geom_label(data = cs2_sel_grid_sub_letters_data[1, ],
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white", size = 8),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color = "black", fill = "white"),
    axis.title = ggplot2::element_blank(),
    legend.position="none") +
  ggplot2::scale_fill_manual(name = "Solution",
                             values = c("discarded" = "#f7fcfd",
                                        "selected" = "#00441b",
                                        "uninhabited" = "grey65")) +
  ggplot2::coord_cartesian(
      xlim = cs2_grid_sub_polygons@bbox[1, ],
      ylim = cs2_grid_sub_polygons@bbox[2, ]) +
  ggplot2::facet_wrap(~ Title1)
p2 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = countries_fpolygons,
                        ggplot2::aes(x=long, y=lat, group=group),
                        fill="grey85", color="grey70") +
  ggplot2::geom_polygon(data = cs2_sel_grid_sub_fpolygons,
                        ggplot2::aes(x=long, y=lat, group=group,
                                     fill = Amount_genetic_selection),
                                     color="black") +
  ggplot2::geom_label(data = cs2_sel_grid_sub_letters_data[2, ],
                      ggplot2::aes(label = letter, x =x , y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white", size = 8),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color = "black", fill = "white"),
    legend.position = "none") +
  ggplot2::scale_fill_manual(name = "Solution",
                             values = c("discarded" = "#f7fcfd",
                                        "selected" = "#00441b",
                                        "uninhabited" = "grey65")) +
  ggplot2::coord_cartesian(
    xlim = cs2_grid_sub_polygons@bbox[1, ],
    ylim = cs2_grid_sub_polygons@bbox[2, ]) +
  ggplot2::facet_wrap(~ Title2)
p3 <- ggplot2::ggplot() +
  ggplot2::geom_point(data = dplyr::arrange(cs2_sel_space_data,
                                            Amount_selection),
                      ggplot2::aes(x = d1, y = d2, fill = Amount_selection),
                      pch = 21, size = 3, color = "black") +
  ggplot2::geom_label(data = cs2_sel_space_letters_data[1,],
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::xlab("NMDS 1") +
  ggplot2::ylab("NMDS 2") +
  ggplot2::scale_x_continuous(expand = c(0.08, 0)) +
  ggplot2::scale_y_continuous(expand = c(0.08, 0)) +
  ggplot2::theme(
      legend.position="none",
      axis.line.x = ggplot2::element_line(),
      axis.line.y = ggplot2::element_line(),
      axis.text = ggplot2::element_text(size = 8),
      axis.title = ggplot2::element_text(size = 8),
      plot.margin = ggplot2::unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
  ggplot2::scale_fill_manual(name = "Solution",
                             values = c("discarded" = "transparent",
                                        "selected" = "#00441b"))
p4 <- ggplot2::ggplot() +
  ggplot2::geom_point(data = dplyr::arrange(cs2_sel_space_data,
                                            Amount_genetic_selection),
                      ggplot2::aes(x = d1, y = d2,
                                   fill = Amount_genetic_selection),
                       pch = 21, size = 3, color = "black") +
  ggplot2::geom_label(data = cs2_sel_space_letters_data[2,],
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size = 3) +
  ggplot2::theme_classic() +
  ggplot2::xlab("NMDS 1") +
  ggplot2::ylab("NMDS 2") +
  ggplot2::scale_x_continuous(expand = c(0.08, 0)) +
  ggplot2::scale_y_continuous(expand = c(0.08, 0)) +
  ggplot2::theme(
    legend.position="none",
    axis.line.x = ggplot2::element_line(),
    axis.line.y = ggplot2::element_line(),
    axis.text = ggplot2::element_text(size = 8),
    axis.title = ggplot2::element_text(size = 8),
    plot.margin = ggplot2::unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
  ggplot2::scale_fill_manual(name = "Solution",
                             values = c("discarded" = "transparent",
                                        "selected" = "#00441b"))
# render plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```