### 2017-04-06 code/rmarkdown/article.tex
r pkgs

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={raptr: Representative and Adequate Prioritization Toolkit in R},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{raptr: Representative and Adequate Prioritization Toolkit in R}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson\(^1\), Jonathan R. Rhodes\(^2\), Hugh P.
Possingham\(^{1,3}\), Richard A. Fuller\(^1\)\\
\(^1\)School of Biological Sciences, The University of Queensland,
Brisbane, QLD, Australia\\
\(^2\)School of Earth and Environmental Sciences, The University of
Queensland, Brisbane, QLD, Australia\\
\(^3\)The Nature Conservancy, South Brisbane, QLD, Australia\\
Correspondence should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{\nolinkurl{jeffrey.hanson@uqconnect.edu.au}}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{06 April 2017}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,lineno,booktabs,subfiles}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% line numbers
\linenumbers

% allow breaks in equations
\allowdisplaybreaks

% format abstract
\renewcommand{\abstractname}{Summary}
\renewenvironment{abstract}
 {\small
  \begin{center}
  \bfseries \abstractname\vspace{-.5em}\vspace{0pt}
  \end{center}
  \list{} {%
   \setlength{\leftmargin}{2mm}
   \setlength{\rightmargin}{\leftmargin}%
  }%
  \item\relax}
{\endlist}

% format section headers
\titleformat*{\section}{\Large\bfseries}
\titleformat{\subsection}[display]
	{\large\sffamily\lsstyle}
	{\subsectiontitlename\ \thesubsection}
	{0.5em}{}
\titleformat*{\subsubsection}{\large\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

% define struts for tables
\newcommand\T{\rule{0pt}{2.6ex}} % top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut

% define command to put new lines in table cells
%\newcommand{\specialcell}[2][c]{%
%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\begin{document}
\maketitle

Abstract word count: 347 / 350\\
Total word count: XXXX / 7000\\
Number of references: 42\\
Number of figures: 8\\
Number of tables: 0\\
Number of textboxes: 0\\
Number of equations: 4\\
Short running title: Short abstract: Prioritize biodiversity processes
and patterns using raptr \url{https://github.com/jeffreyhanson/raptr}\\
Keywords: conservation, biodiversity, mixed integer linear programming,
optimization, protected areas

\clearpage

\section{Summary}\label{summary}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  An underlying aim in conservation planning is to maximize the
  long-term persistence of biodiversity. To fulfill this aim, the
  ecological and evolutionary processes that sustain biodiversity must
  be preserved. One way to conserve such processes at the feature level
  (eg. species, ecosystem) is to preserve a sample of the feature (eg.
  individuals, areas) that is representative in terms of the intrinsic
  or extrinsic physical attributes that underpin the process of
  interest. For example, by conserving a sample of populations with
  local adaptations that is representative in terms of the full
  variation found in the species--physical attributes associated with
  adaptation--protected areas can foster adaptive processes by ensuring
  these adaptations are not lost. Despite this, current reserve
  selection methods overwhelmingly focus on securing an adequate amount
  of area or habitat for each feature. Little attention has been
  directed towards capturing a representative sample of each feature.
\item
  To address this issue, we developed the \texttt{raptr R} package to
  help guide reserve selection. Users set ``amount targets''--similar to
  conventional methods--to ensure that solutions secure a sufficient
  proportion of area or habitat for each feature. Additionally, users
  set ``space targets'' to secure a representative sample of variation
  in ecologically or evolutionary relevant attributes (eg. environmental
  or genetic variation). We demonstrate the functionality of this
  package using simulations and two case studies. In these studies, we
  generated solutions using amount targets--representing conventional
  methods--and compared them with solutions generated using amount and
  space targets.
\item
  Our results demonstrate that different solutions emerge when targeting
  a representative sample of each feature. We show that using these
  targets is important for features that have multimodal distributions
  in the process-related attributes (eg. species with multimodal
  niches). We also found that solutions could conserve a far more
  representative sample with only a slight increase in reserve size.
\item
  The \texttt{raptr R} package provides a toolkit for making
  prioritizations that secure an adequate and representative sample of
  each feature. By using solutions that secure a representative sample
  of each feature, prioritizations may have a greater chance of
  achieving long-term biodiversity persistence.
\end{enumerate}

\clearpage
\input{./text.tex}

\clearpage
\input{./figures.tex}

\end{document}